---
title: "Continuous RVs and Universality of Uniform"
author: "Your name"
date: "2025-04-04"
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 3)
```

## Continuous RVs

Built into `R` are some functions that work with the named continuous distributions. Recall we have the following types of functions:

For example:

-   `dnorm(x, mu, sigma)` will evaluate the PDF of a $N(\mu, \sigma^2)$ distribution at the value `x`.
-   `pnorm(q, mu, sigma)` will evaluate the CDF of a $N(\mu, \sigma^2)$ distribution at the value `q`
-   `qnorm(p, mu, sigma)` will evaluate the inverse-CDF of a $N(\mu, \sigma^2)$ distribution at the value `p`
-   `rnorm(n, mu, sigma)` will generate `n` random variables from the $N(\mu, \sigma^2)$ distribution

NOTE: for the `_norm()` functions in R, the functions expect standard deviation as input, not variance!!

```{r}
# X ~ N(0, 4) -> What is P(X <= 1):

# Generate 10 random variables from the Unif(0,1) distribution:

# X ~ Exp(2). What is f(1)?

# Obtain median of standard normal

```

## Visualizing densities

Sometimes it can be helpful to visualize the density of a distribution. There are a couple ways we can do this. Let's do this example for the $Unif(1, 1.5)$ distribution.

### Option 1

*If we can randomly sample from the distribution*, we can generate lots and lots of random variables from that distribution and make a histogram of them:

```{r fig.height=3}
# simulate lots and lots of Unif(1, 1.5) rvs

# make a histogram of the simulations
```

### Option 2

*If we have access to the PDF directly* (which we do for the named distributions), we can simply graph the function:

```{r fig.height = 3}
# create a sequence of values from 0 and 2 at equally-size 0.01 increments


# evaluate PDF of interest at each value in x_seq


# visualize


```

## Universality of Uniform / Probability Integral Transform

Now, let's see the University of the Uniform in action! Suppose you've lost access to all the functions in R that allow you randomly generate rvs from all the named distribution except for the Uniform.

### Example 1

How can we simulate values from the $\text{Exp}(2)$ distribution?

```{r fig.height = 3}
# generate lots of Unif(0,1) rvs

# use inverse CDF that we derived

# visualize 

```

### Example 2

Suppose we have a distribution whose PDF is $$f_{X}(x) = \frac{e^{-x}}{(1 + e^{-x})^2}$$

for $x \in (-\infty, \infty)$.

Write code to simulate 1000 random variables from this distribution, and and visualize them as a density histogram.

```{r fig.height = 3}


```
