---
title: "Intro to R and Simulations"
author: "Your name here"
date: "02-19-2025"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

## R is a calculator 

```{r}

```

R respects PEMDAS:

```{r}

```

### Booleans

Booleans/logicals can be formed with (in)equality symbols:

```{r}

```

### Variables 

Storing/saving values into variables is achieved by using the syntax `<variable name> <- <value>`. Note that the variable name cannot begin with a number! The result of a stored variable is not displayed to the user. Type out the variable name after storing to explicitly show the values.

```{r}

```

### Vectors 

Create vectors of values (i.e. a list of values) using the concatenate `c()` function:

```{r}

```

The syntax `a:b` creates a sequence of integers from `a` to `b`.

```{r}

```

R is vectorized language, so it operates element-wise in vectors. Also, under the hood, booleans/logicals TRUE/FALSE are treated as 1/0:

```{r}

```

## Functions

A lot of the commands we have been working with are functions. You can tell a command is a function if it has parentheses (e.g. sum() and c()). A function usually takes in some argument(s) as input and returns something back as output. We can write custom functions using the function `function()`. We need to specify the name of the function, what it expects as input, as well as what to output.

### Birthday problem, generalized

Suppose a group of $k$ people each choose a number randomly with replacement from a list of $n$ distinct numbers. The probability
of at least one match in the group is 1 minus the probability of no matches:

$$1 - \binom{n}{k} \frac{k!}{n^k}$$

We can write this as a function in R:

```{r}


```

Note that the function name is prob_match, and the arguments/inputs are called `n` and `k`. The output is the probability of at least one match for the specified values of `n` and `k`. We can now use this function. For example, the probability of at least one match among 3 people choosing a number between 1 and 40 is:

```{r}


```

## for looping

Suppose I want to obtain the probability of a match for $n = 40$ distinct numbers among a range of $k = 2, \ldots, 25$ people. Using what's called a "for loop" can help automate this process:

```{r}


```

## Basic plotting

We can plot one variable against another (e.g. input and output of a function) using the `plot()` function. The
first argument is the x-axis variable (e.g. input), and the second argument is the y-axis variable (e.g. output).

For example, we will plot probabilities of each least one match we obtained above: 


```{r fig.height=4}


```



## Sampling and simulating

The `sample()` function generates random values from a pool of values contained in a specific vector. You
can twist many knobs to get sampling with or without replacement, uneven probabilities, and different sized
samples.

```{r}


```

## Approximating probabilities

In the following, simulation flipping a fair coin once:

```{r}


```

We can approximate probabilities by counting up repeatedly simulating an experiment, counting up how many times a favorable outcome occurred, and dividing by the total number of experiments we did. This is easily done using the `replicate()` function:

```{r}


```
